---
title: "Explanations in natural language"
author: "Adam Izdebski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Describing-Explanations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  fig.width = 7,
  fig.height = 3.5,
  warning = FALSE,
  message = FALSE
)
```


# Introduction

We adress the problem of insuficient interpretability of explanations for domain experts. We solve this issue by introducing `describe()` function, which automaticly generates natural language descriptions of explanations generated with `ingredients` package.

# ingredients Package

The `ingredients` package allows for generating prediction validation and predition perturbation explanations. They allow for both global and local model explanation.

Generic function `decribe()` generates a natural language description for explanations generated with `feature_importance()`, `ceteris_paribus()` functions.

To show generating automatic descriptions we first load the data set and build a random forest model classifying, which of the passangers survived sinking of the titanic. Then, using `DALEX` package, we generate an explainer of the model. Lastly we select a random passanger, which prediction's should be explained.

```{r message=FALSE, warning=FALSE}
library("DALEX")
library("ingredients")
library("randomForest")
titanic <- na.omit(titanic)

model_titanic_rf <- randomForest(survived == "yes" ~ .,
                                 data = titanic)

explain_titanic_rf <- explain(model_titanic_rf,
                            data = titanic[,-9],
                            y = titanic$survived == "yes",
                            label = "Random Forest")

passanger <- titanic[sample(nrow(titanic), 1) ,-9]
passanger
```

Now we are ready for generating various explantions and then describing it with `describe()` function.

## Feature Importance

Feature importance explanation shows the importance of all the model's variables. As it is a global explanation technique, no passanger need to be specified.

```{r}
importance_rf <- feature_importance(explain_titanic_rf)
plot(importance_rf)
```

Function `describe()` easily describes which variables are the most important.
Argument `nonsignificance_treshold` as always sets the level above which variables become significant. For higher treshold, less variables will be described as significant.  

```{r}
describe(importance_rf)
```

## Ceteris Paribus Profiles

Ceteris Paribus profiles shows how the model's input changes with the change of a specified variable.

```{r}
perturbed_variable <- "class"
cp_rf <- ceteris_paribus(explain_titanic_rf,
                         passanger,
                         variables = perturbed_variable)
plot(cp_rf, variable_type = "categorical")
```

For a user with no experience, interpreting the above plot may be not straightforward. Thus we generate a natural language description in order to make it easier.

```{r}
describe(cp_rf)
```

Natural lannguage descriptions should be flexible in order to provide the desired level of complexity and specificity. Thus various parameters can modify the description being generated.

```{r}
describe(cp_rf,
         display_numbers = TRUE,
         label = "the probability that the passanger will survive")
```

Please note, that `describe()` can handle only one variable at a time, so it is recommended to specify, which variables should be described.

```{r}
describe(cp_rf,
         display_numbers = TRUE,
         label = "the probability that the passanger will survive",
         variables = perturbed_variable)
```

Continuous variables are described as well.

```{r}
perturbed_variable_continuous <- "age"
cp_rf <- ceteris_paribus(explain_titanic_rf,
                         passanger)
plot(cp_rf, variables = perturbed_variable_continuous)
describe(cp_rf, variables = perturbed_variable_continuous)
```

Ceteris Paribus profiles are described only for a single observation. If we want to access the influence of more than one observation, we need to describe dependency profiles.

## Partial Dependency Profiles

```{r}
pdp <- aggregate_profiles(cp_rf, type = "partial")
plot(pdp, variables = "fare")
describe(pdp, variables = "fare")
```

```{r}
pdp <- aggregate_profiles(cp_rf, type = "partial", variable_type = "categorical")
plot(pdp, variables = perturbed_variable)
describe(pdp, variables = perturbed_variable)
```
